% This file is part of the Weak Lensing project.
% Copyright 2013 David W. Hogg (NYU).

\documentclass[12pt, pdftex]{article}

\newcommand{\dd}{\mathrm{d}}
\newcommand{\given}{\,|\,}
\newcommand{\RA}{\mathrm{RA}}
\newcommand{\Dec}{\mathrm{Dec}}
\newcommand{\shearpars}{\theta}
\newcommand{\position}{x}
\newcommand{\cosmopars}{\omega}
\newcommand{\densitypars}{\rho}
\newcommand{\data}{D}
\newcommand{\shape}{s}
\newcommand{\shapepars}{\alpha}

\begin{document}

\section*{white paper: Probabilistically righteous galaxy--galaxy weak lensing.}

\noindent
\textsl{[by Hogg, Bolton, Marshall, others]}

\noindent
\textsl{2013-02-12}

In galaxy--galaxy weak lensing, standard operating procedure is to
measure the ellipticities of large numbers of background galaxies and
average their ellipticity vectors (or spinors) in annuli around target
foreground galaxies.  The idea is to choose foreground galaxies that
are expected to be similar in their lensing or mass properties, or
galaxies whose expected properties can be scaled to be similar, and to
do the averaging of the vectors in polar coordinates.  If there is a
weak-lensing signal, then the mean ellipticity will not be zero.
Background galaxies will be on average tangentially elongated, and the
mean ellipticity will be a function of radius that depends (very
simply) on the mean projected mass distribution in the foreground
galaxies, or really the mean projected excess mass \emph{correlated
  with} the foreground galaxies.

In general, because the signals are weak and the background galaxy
density is not high, data from the vicinities of many foreground
galaxies must be averaged together to get good signal-to-noise.  Any
time ``stacking'' or ``averaging'' is taking place, there is usually a
``modeling'' alternative that is more principled, more flexible,
better able to account for odd noise properties, and more justifiable
in terms of probabilistic inference.  In the case of galaxy--galaxy
weak lensing, not only is this true, it is certainly the case that the
noise properties are odd: The noise in the measurement comes from
``shape noise'' or really a combination of the fact that background
galaxies---even unlensed---have a non-trivial shape distribution and
also that they are not measured properly.

When we write this document, it may or may not include the following
equations:\ldots

\end{document}
