% This file is part of the Weak Lensing project.
% Copyright 2011, 2012 David W. Hogg (NYU) and Philip J. Marshall (Oxford).

\documentclass[12pt]{article}
\begin{document}

\section*{Hierarchical inference of cosmic shear maps}
\noindent
David~W.~Hogg \\
\textsl{Center for Cosmology and Particle Physics, New York University} \\
and \textsl{Max-Planck-Institut f\"ur Astronomie} \\[1ex]
Phillip~J.~Marshall \\
\textsl{Oxford University}

\begin{abstract}
We set up a toy weak-lensing problem---an overly simple scenario in
which the two-dimensional morphologies of galaxies under study
can be completely described by two components of a complex ellipticity---and 
use it to look at the
methods by which a shear map can be inferred.  While it is true that
the average ellipticity of a sufficiently large sample of observed
galaxy images does provide an estimate of the cosmological shear, we
show that it is more precise and accurate to probabilistically
infer the shear map from the distribution of observations.  This
inference is hierarchical in that it involves learning the parameters
of an exceedingly flexible model for the distribution of unlensed
morphologies, subject to a constraint that this unlensed distribution
has no mean ellipticity.  Although the problem we solve here is a toy,
all real issues with real data are expected to amplify the relative
value of hierarchical probabilistic inference over brute sample
averaging.  We make no representations, however, about computational
costs.
\end{abstract}

...content here...

Introduction argument: Imagine (unrealistically) that all galaxies
have the same non-trivial non-circular shape, and are all exemplars of
that same shape, but with different orientations on the sky.  Then
even the observation of a \emph{single} galaxy would provide a
significant local measure of the shear map.  Taking ellipticities and
averaging would be wasteful and imprecise relative to two-dimensional
morphological modeling.  For example, what if all unlensed galaxies in
some color-selected sample are precisely two-dimensional Gaussians
that are 2 arcsec along the major axis and 1 arcsec along the minor
axis.  That's too simple for reality---Galaxies live in
three-dimensional space, after all, so there are two other Euler
angles---but it shows that averaging can never capitalize on
informative features in the distribution of morphologies.

Another introduction argument: Show (if it is indeed possible to do
so) that if the distribution of morphologies is ``featureless'' in the
relevant ways, hierarchical inference reduces to averaging.  Is that
true or possible?

Define terms, especially ``What constitutes a \emph{shear map}?''

\section{A toy hierarchical model}

Suppose we have measured a set of galaxy shapes in the presence of noise and
PSF blurring, but have somehow managed to do this accurately and such that
their shape information can be fully captured by their complex ellipticities,
and the uncertainties on their components (which are Gaussian distributed).
The next task is to infer the shear field that has been applied to those
shapes---suppose that shear field is constant, and small over a small patch of
sky. The measured ellipticity of a galaxy $\epsilon_{\rm obs}$ 
is related to its intrinsic
ellipticity $\epsilon$ and the applied (reduced) shear $g$ by 
\begin{equation}
\epsilon_{\rm obs} = \epsilon + g + n.
\end{equation}
If the distributions for the $\epsilon$ and the noise $n$ have zero mean, then
an unbiased estimator for $g$ is the ensemble average of $\epsilon_{\rm obs}$
in the patch, and its variance is the sum of the variances of $\epsilon$ and
$n$. However, taking this average ignores all the information we have
about the distribution for $\epsilon$ that is contained in all the data we
have, in this patch and in all others: by inferring the PDF $p(\epsilon)$ at
the same time as $g$, we hope to effectively reduce the variance on $g$. This
reduction would be most obvious if there was a sub-population of sources that
had smaller intrinsic ellipticities. These could be marked out by their red
colors, for example.

\section{Tests}

\section{Results}

\section{Discussion}

\end{document}
